The yaq Project: \\ Standardized Software Enabling Flexible Instrumentation

### 2022-11-11 paper.tex
minor changes (#29)
* minor changes

* Update paper.tex

Co-authored-by: Blaise Thompson <blaise@untzag.com>
hardware recount

The Wright Group at UW-Madison needs to orchestrate a large variety of hardware in multidimensional scans for their complex spectroscopy experiments \cite{MukamelShaul2000a, WrightJohnCurtis2011a}.
This need for exquisite hardware control has resulted in several prior attempts at ``home-built'' orchestration software \cite{CarlsonRogerJohn1988a, MeyerKentAlbert2004b, KainSchuyler2017a, ThompsonBlaiseJonathan2018a, SundenKyle2022}.
Now, using \yaq{}, the Wright Group has been able to move to Bluesky rather than inventing their own sophisticated control software ``from scratch''.
The Wright Group uses simulated hardware to enable client development away from the active laboratory computers.
This allows Wright Group researchers to create polished client interfaces without interrupting  ongoing experiments.
Clients developed by the Wright Group have proved flexible enough to be used on four laser systems, each with different complements of hardware.
The Wright Group expects to spend less energy developing control software and more energy developing creative spectroscopy experiments in future work.

@@

Figure \ref{fig:histogram} shows the distribution of unique lines of Python code written to implement each of the daemons in our current ecosystem.
While lines of code is an imperfect metric, we use it here to represent the amount of work required to create a daemon for a new hardware peripheral.
Most interfaces, such as Brooks MFC\footnote{\texttt{yaqd-brooks-mfc-gf} source code available on GitHub at \texttt{https://github.com/yaq-project/yaqd-brooks/}}, have been implemented in fewer than 100 lines of Python code.
Even the most complicated daemons are implemented in about 300 lines.
Implementing \yaq{} daemons using Python is enabled through our own tooling mentioned above and the large and growing ecosystem of hardware interface tooling libraries that Python now provides\cite{pyserial, pyusb, pyvisa, pymodbus}.
In our experience, the process of creating a new daemon involves about a day of work after communication with the hardware using Python is well understood.
There are currently 70 daemons in the yaq project supporting at least 50 kinds of hardware, noting that some daemons support the same hardware and others are software only.
Because \yaq{} is standards based, anyone can design and publish new daemons extending our hardware support.
A living list of all daemons and supported hardware can be found on the \yaq{} website.\footnote{See https://yaq.fyi/daemons/ and https://yaq.fyi/hardware/ for living lists of all \yaq{} daemons and supported hardware, respectively.}

>>>>

The Wright Group at UW-Madison needs to orchestrate a large variety of hardware in multidimensional scans for their complex spectroscopy experiments \cite{MukamelShaul2000a, WrightJohnCurtis2011a}.
This need for exquisite hardware control has resulted in several prior attempts at ``home-built'' orchestration software \cite{CarlsonRogerJohn1988a, MeyerKentAlbert2004b, KainSchuyler2017a, ThompsonBlaiseJonathan2018a, SundenKyle2022}.
Now, using \yaq{}, the Wright Group has been able to move to Bluesky rather than inventing their own sophisticated control software ``from scratch''.
The Wright Group uses simulated hardware to enable client development away from the active laboratory computers.
This allows Wright Group researchers to create polished client interfaces without interrupting  ongoing experiments.
Clients developed by the Wright Group have proved flexible enough to be used on four laser systems, each with different complements of hardware.
Moving forward, the Wright Group will spend less energy developing control software and more energy developing creative spectroscopy experiments.

@@

Figure \ref{fig:histogram} shows the distribution of unique lines of Python code written to implement each of the daemons in our current ecosystem.
While lines of code is an imperfect metric, we use it here to represent the amount of work required to create a daemon for a new hardware peripheral.
Most interfaces, such as Brooks MFC\footnote{\texttt{yaqd-brooks-mfc-gf} source code available on GitHub at \texttt{https://github.com/yaq-project/yaqd-brooks/}}, have been implemented in fewer than 100 lines of Python code.
Even the most complicated daemons are implemented in about 300 lines.
Implementing \yaq{} daemons using Python is enabled through our own tooling mentioned above and the large and growing ecosystem of hardware interface tooling libraries that Python now provides\cite{pyserial, pyusb, pyvisa, pymodbus}.
In our experience, the process of creating a new daemon involves about a day of work after communication with the hardware using Python is well understood.
There are currently 72 daemons in the yaq project supporting at least 66 types of hardware, noting that some daemons support the same hardware and others are software only.
Because \yaq{} is standards based, anyone can design and publish new daemons extending our hardware support.
A living list of all daemons and supported hardware can be found on the \yaq{} website.\footnote{See https://yaq.fyi/daemons/ and https://yaq.fyi/hardware/ for living lists of all \yaq{} daemons and supported hardware, respectively.}
